Primary analysis

### 2022-09-15 analysis/primary.Rmd
update primary analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

# Outline
- IOIs
- KOT
- KV
- KV-Diff

@@

```{r file, include = FALSE}
filename = "filtered/dt_correct_onset_1.txt"
```

```{r extract, include = FALSE}
data <- fread(filename, header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
data$SubNr <- as.factor(data$SubNr)

# assign stimuli categories
data$Stimuli <- as.numeric(gsub(".mid", "", data$Stimuli))
data$Category <- "NA"
data[Stimuli < 5]$Category <- "art_only"
data[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
data[Stimuli > 8 & Stimuli < 13]$Category <- "high"
data[Stimuli > 12]$Category <- "low"

# assign articulation (yes/no) & dynamics (yes/no)
data$Articulation <- "NA"
data$Dynamics <- "NA"
data[Category == "art_only"]$Articulation <- "Present"
data[Category == "art_only"]$Dynamics <- "Absent"
data[Category == "dyn_only"]$Articulation <- "Absent"
data[Category == "dyn_only"]$Dynamics <- "Present"
data[Category == "high"]$Articulation <- "Present"
data[Category == "high"]$Dynamics <- "Present"
data[Category == "low"]$Articulation <- "Absent"
data[Category == "low"]$Dynamics <- "Absent"

# IOIs
dt_ioi <- data
dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))

# remove the first note
dt_ioi <- dt_ioi[RowNr != 1]

# assign Interval
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# define intervals
ls_l_f <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_s_p <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
change_1 <- c(5, 17, 47) # legato to staccato / forte to piano
change_2 <- c(8, 20, 39) # staccato to legato / piano to forte

# assign Subcomponents
dt_ioi$Subcomponent <- "NA"
# legato & forte
for (phrase in 1:length(ls_l_f)){
  for (note in 1:length(ls_l_f[[phrase]])){
    dt_ioi[Interval == ls_l_f[[phrase]][note]]$Subcomponent <- "L-F"
  }
}
# staccato & piano
for (phrase in 1:length(ls_s_p)){
  for (note in 1:length(ls_s_p[[phrase]])){
    dt_ioi[Interval == ls_s_p[[phrase]][note]]$Subcomponent <- "S-P"
  }
}
for (number in change_1){
  dt_ioi[Interval == number]$Subcomponent <- "LFtoPS"
}
for (number in change_2){
  dt_ioi[Interval == number]$Subcomponent <- "SPtoLF"
}

# KOT

# KV
dt_vel <- data
dt_vel$Diff <- diff(c(0, dt_vel$Velocity))

# remove the first note
dt_vel_diff <- dt_vel[RowNr != 1]
dt_vel$Diff <- NULL # remove Diff from dt_vel

# assign Interval
dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)

# for each note (velocity)
ls_l_f_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_s_p_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# assign Subcomponents
# for each note
dt_vel$Subcomponent <- "NA"
# legato & forte
for (phrase in 1:length(ls_l_f_2)){
  for (note in 1:length(ls_l_f_2[[phrase]])){
    dt_vel[RowNr == ls_l_f_2[[phrase]][note]]$Subcomponent <- "L-F"
  }
}
# staccato & piano
for (phrase in 1:length(ls_s_p_2)){
  for (note in 1:length(ls_s_p_2[[phrase]])){
    dt_vel[RowNr == ls_s_p_2[[phrase]][note]]$Subcomponent <- "S-P"
  }
}

dt_vel_diff$Subcomponent <- "NA"
# legato & forte
for (phrase in 1:length(ls_l_f)){
  for (note in 1:length(ls_l_f[[phrase]])){
    dt_vel_diff[Interval == ls_l_f[[phrase]][note]]$Subcomponent <- "L-F"
  }
}
# staccato & piano
for (phrase in 1:length(ls_s_p)){
  for (note in 1:length(ls_s_p[[phrase]])){
    dt_vel_diff[Interval == ls_s_p[[phrase]][note]]$Subcomponent <- "S-P"
  }
}
for (number in change_1){
  dt_vel_diff[Interval == number]$Subcomponent <- "LFtoSP"
}
for (number in change_2){
  dt_vel_diff[Interval == number]$Subcomponent <- "SPtoLF"
}
```

@@

```{r ioi, echo = FALSE}
# for each individual
ioi <- dt_ioi[Subcomponent != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Stimuli, Articulation, Dynamics)]
ioi <- ioi[order(SubNr, Stimuli, Articulation, Dynamics)]
ioi
```

@@

```{r ioi-plot, echo = FALSE, fig.width = 4.5}
ggboxplot(ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

@@

```{r ioi-all, echo = FALSE}
# group mean
ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics)]
ioi_all <- ioi_all[order(Articulation, Dynamics)]
ioi_all
```

@@

```{r ioi-all-plot, echo = FALSE}
ggboxplot(ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

# Dynamics
## 1. Individual
```{r vel, echo = FALSE}
# for each individual
vel <- dt_vel[Subcomponent != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
vel <- vel[order(SubNr, Stimuli, Articulation, Dynamics)]
vel
```

@@

```{r vel-plot,  echo = FALSE, fig.width = 4.5}
ggboxplot(vel[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Forte)")
ggboxplot(vel[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Piano)")
```

## 2. All - group
```{r vel-all, echo = FALSE}
# group mean
vel_all <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
vel_all <- vel_all[order(Articulation, Dynamics, Subcomponent)]
vel_all
```

@@

```{r vel-all-plot,  echo = FALSE}
ggboxplot(vel[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
ggboxplot(vel[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
```

# KV-Diff
## 1. Individual

```{r vel-diff, echo = FALSE}
# for each individual
vel_diff <- dt_vel_diff[Subcomponent == "LFtoSP" | Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
vel_diff <- vel_diff[order(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
vel_diff
```

@@

```{r vel-diff-plot, echo = FALSE, fig.width = 4.5}
ggboxplot(vel_diff[Subcomponent == "LFtoSP", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (LFtoSP)")
ggboxplot(vel_diff[Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (SPtoLF)")
```

## 2. All - group
```{r vel-diff-all, echo = FALSE}
# group mean
vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
vel_diff_all <- vel_diff_all[order(Articulation, Dynamics, Subcomponent)]
vel_diff_all
```

@@

```{r vel-diff-all-plot,  echo = FALSE}
ggboxplot(vel_diff[Subcomponent == "LFtoSP", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (LF to SP)")
ggboxplot(vel_diff[Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (SP to LF)")
```

>>>>

@@

```{r file, include = FALSE}
data_ioi = "preprosessor/trimmed/dt_ioi_trimmed.txt"
data_kot = "preprosessor/trimmed/dt_kot_trimmed.txt"
data_vel = "preprosessor/trimmed/dt_vel_trimmed.txt"
data_vel_diff = "preprosessor/trimmed/dt_vel_diff_trimmed.txt"
```

```{r prep, include = FALSE}
dt_ioi <- fread(data_ioi, header = T, sep = ",", dec = ".")
dt_kot <- fread(data_kot, header = T, sep = ",", dec = ".")
dt_vel <- fread(data_vel, header = T, sep = ",", dec = ".")
dt_vel_diff <- fread(data_vel_diff, header = T, sep = ",", dec = ".")

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)

# assign stimuli categories
dt_ioi$Stimuli <- as.numeric(gsub(".mid", "", dt_ioi$Stimuli))
dt_ioi$Articulation2 <- "NA"
dt_ioi$Dynamics2 <- "NA"
dt_ioi[Stimuli < 5]$Articulation2 <- "Present"
dt_ioi[Stimuli < 5]$Dynamics2 <- "Absent"
dt_ioi[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
dt_ioi[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
dt_ioi[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
dt_ioi[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
dt_ioi[Stimuli > 12]$Articulation2 <- "Absent"
dt_ioi[Stimuli > 12]$Dynamics2 <- "Absent"

dt_kot$Stimuli <- as.numeric(gsub(".mid", "", dt_kot$Stimuli))
dt_kot$Articulation2 <- "NA"
dt_kot$Dynamics2 <- "NA"
dt_kot[Stimuli < 5]$Articulation2 <- "Present"
dt_kot[Stimuli < 5]$Dynamics2 <- "Absent"
dt_kot[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
dt_kot[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
dt_kot[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
dt_kot[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
dt_kot[Stimuli > 12]$Articulation2 <- "Absent"
dt_kot[Stimuli > 12]$Dynamics2 <- "Absent"

dt_vel$Stimuli <- as.numeric(gsub(".mid", "", dt_vel$Stimuli))
dt_vel$Articulation2 <- "NA"
dt_vel$Dynamics2 <- "NA"
dt_vel[Stimuli < 5]$Articulation2 <- "Present"
dt_vel[Stimuli < 5]$Dynamics2 <- "Absent"
dt_vel[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
dt_vel[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
dt_vel[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
dt_vel[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
dt_vel[Stimuli > 12]$Articulation2 <- "Absent"
dt_vel[Stimuli > 12]$Dynamics2 <- "Absent"

dt_vel_diff$Stimuli <- as.numeric(gsub(".mid", "", dt_vel_diff$Stimuli))
dt_vel_diff$Articulation2 <- "NA"
dt_vel_diff$Dynamics2 <- "NA"
dt_vel_diff[Stimuli < 5]$Articulation2 <- "Present"
dt_vel_diff[Stimuli < 5]$Dynamics2 <- "Absent"
dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
dt_vel_diff[Stimuli > 12]$Articulation2 <- "Absent"
dt_vel_diff[Stimuli > 12]$Dynamics2 <- "Absent"
```

@@

```{r ioi, echo = FALSE}
# for each individual
ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2)]

ioi <- ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2)]
ioi
```

@@

```{r ioi-plot, echo = FALSE, fig.width = 4.5}
ggboxplot(ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation2, Dynamics2)], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

@@

```{r ioi-all, echo = FALSE}
# group mean
ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2)]
ioi_all
```

@@

```{r ioi-all-plot, echo = FALSE}
ggboxplot(ioi, "Articulation2", "Mean", color = "Dynamics2", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

# Articulation
## 1. Individual
```{r kot, echo = FALSE}
# for each individual
kot_trial <- dt_kot[Articulation == "Legato" | Articulation == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Articulation)]

kot <- kot_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2, Articulation)]
kot
```

@@

```{r kot-plot,  echo = FALSE, fig.width = 4.5}
ggboxplot(kot_trial[Articulation == "Legato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")

ggboxplot(kot_trial[Articulation == "Staccato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
```

## 2. All - group
```{r kot-all, echo = FALSE}
# group mean
kot_all <- kot[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2, Articulation)]
kot_all
```

@@

```{r kot-all-plot, echo = FALSE}
ggboxplot(kot[Articulation == "Legato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")

ggboxplot(kot[Articulation == "Staccato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
```

# Dynamics
## 1. Individual
```{r vel, echo = FALSE}
# for each individual
vel_trial <- dt_vel[Dynamics != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Dynamics)]

vel <- vel_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2, Dynamics)]
vel
```

@@

```{r vel-plot, echo = FALSE, fig.width = 4.5}
ggboxplot(vel_trial[Dynamics == "Forte"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Forte)")

ggboxplot(vel_trial[Dynamics == "Piano"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Piano)")
```

## 2. All - group
```{r vel-all, echo = FALSE}
# group mean
vel_all <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2, Dynamics)]
vel_all
```

@@

```{r vel-all-plot,  echo = FALSE}
ggboxplot(vel[Dynamics == "Forte"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")

ggboxplot(vel[Dynamics == "Piano"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
```

# Dynamics Difference
## 1. Individual
```{r vel-diff, echo = FALSE}
# for each individual
vel_diff_trial <- dt_vel_diff[Dynamics == "FtoP" | Dynamics == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Dynamics)]

vel_diff <- vel_diff_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2, Dynamics)]
vel_diff
```

```{r vel-diff-plot, echo = FALSE, fig.width = 4.5}
ggboxplot(vel_diff_trial[Dynamics == "FtoP"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (FtoP)")

ggboxplot(vel_diff_trial[Dynamics == "PtoF"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (PtoF)")
```

## 2. All - group
```{r vel-diff-all, echo = FALSE}
# group mean
vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2, Dynamics)]
vel_diff_all
```

```{r vel-diff-all-plot,  echo = FALSE}
ggboxplot(vel_diff[Dynamics == "FtoP"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (FtoP)")

ggboxplot(vel_diff[Dynamics == "PtoF"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (PtoF)")
```
